\section{Value Encoding}

The \module{hla.omt} module provides
\begin{itemize}
\item Automatic IEEE1516.2 value encoding (pack and unpack functions).
\item Direct import of OMT DIF files.
\end{itemize}

First, import the module functions
\begin{verbatim}
import hla.omt as fom
\end{verbatim}

Each encoding object provides the \method{pack}, \method{unpack} methods, and
the \member{octetBoundary} property.

\begin{classdesc}{HLAbasic}{}
\begin{methoddesc}[HLAbasic]{pack}{value}
Returns encoded value.
\begin{verbatim}
encodedValue = fom.HLAfloat64LE.pack(3.14)
encodedValue += fom.HLAinteger16BE.pack(42)
\end{verbatim}
\end{methoddesc}
\begin{methoddesc}[HLAbasic]{unpack}{value}
Returns a (value, offset) tuple. The offset indicates number of consumed
bytes, i.e. offset of the next unpack operation.
\begin{verbatim}
value1, offset2 = fom.HLAfloat64LE.unpack(encodedValue)
value2, offset3 = fom.HLAinteger16BE.unpack(encodedValue[offset2:])
print value1, value2
\end{verbatim}
\end{methoddesc}
\begin{memberdesc}[int]{octetBoundary}
Read-only property. The octet boundary (in bytes) of this data type.
\end{memberdesc}
\end{classdesc}

The \module{hla.omt} module defines the following basic types.
\begin{tableiii}{l|l|l}{constant}{type name}{bytes}{endian}
\lineiii{HLAinteger16BE}{2}{Big}
\lineiii{HLAinteger32BE}{4}{Big}
\lineiii{HLAinteger64BE}{8}{Big}
\lineiii{HLAfloat32BE}{4}{Big}
\lineiii{HLAfloat64BE}{8}{Big}
\lineiii{HLAinteger16LE}{2}{Little}
\lineiii{HLAinteger32LE}{4}{Little}
\lineiii{HLAinteger64LE}{8}{Little}
\lineiii{HLAfloat32LE}{4}{Little}
\lineiii{HLAfloat64LE}{8}{Little}
\end{tableiii}

Additional data types can be defined using the \class{HLAencoding},
\class{HLAfixedArray}, \class{HLAvariableArray}, or \class{HLAfixedRecord}
class, or imported from an OMT DIF file.

\begin{funcdesc}{HLAuse}{filename}
Imports data types from an OMT DIF file. Format of the OMT DIF file is
defined in IEEE 1516.2.

All the RPR-FOM classes and data types will become accessible from the
module's dictionary.

\begin{verbatim}
fom.HLAuse('rpr2-d18.xml')

encodedValue = fom.WorldLocationStruct.pack({'X':0, 'Y':3.12, 'Z':1})
value, offset = fom.WorldLocationStruct.unpack(encodedValue)
\end{verbatim}

\end{funcdesc}

\begin{classdesc}{HLAencoding}{typename}
Facilitates a forward type declaration. The typename is resolved from the
module's dictionary once the \method{pack} or \method{unpack} method is called.

\begin{verbatim}
newType = fom.HLAencoding('WorldLocationStruct')
\end{verbatim}
is equivalent to the following OMT DIF
\begin{verbatim}
<simpleData name="newType" representation="WorldLocationStruct"/>
\end{verbatim}
\end{classdesc}

\begin{classdesc}{HLAenumerated}{name, representation, \{enumeratorName:value\}}
Defines an enumeration.

\begin{verbatim}
HLAboolean = fom.HLAenumerated("HLAboolean",
    HLAinteger32BE, {"HLAfalse":0, "HLAtrue":1})
\end{verbatim}
is equivalent to the following OMT DIF
\begin{verbatim}
<enumeratedData name="HLAboolean" representation="HLAinteger32BE">
    <enumerator name="HLAfalse" values="0"/>
    <enumerator name="HLAtrue" values="1"/>
</enumeratedData>
\end{verbatim}

The enumerators are accessible via attributes. The values behave like integers,
but don't support arithmetic.
\begin{verbatim}
print HLAboolean.HLAtrue
\end{verbatim}

The \method{pack} function takes an enumerator, or an integer value.
\begin{verbatim}
encodedValue = fom.HLAboolean.pack(HLAboolean.HLAtrue)
\end{verbatim}
\end{classdesc}

\begin{classdesc}{HLAfixedArray}{name, elementType, cardinality}
Defines a fixed-length array.

\begin{verbatim}
arrayType = fom.HLAfixedArray("arrayType", fom.HLAfloat32LE, 3)
\end{verbatim}
is equivalent to the following OMT DIF
\begin{verbatim}
<arrayData name="arrayType"
    encoding="HLAfixedArray" dataType="HLAfloat32LE" cardinality="3"/>
\end{verbatim}

The \method{pack} function takes a sequence object.
\begin{verbatim}
encodedValue = fom.arrayType.pack([3.14, 17, 42])
\end{verbatim}
\end{classdesc}

\begin{classdesc}{HLAvariableArray}{name, elementType, cardinality = None}
Defines a variable-length array.

Dynamic cardinality is represented by the \class{None} value.

\begin{verbatim}
newType = fom.HLAvariableArray("newType", fom.HLAinteger16BE, None)
\end{verbatim}
is equivalent to the following OMT DIF
\begin{verbatim}
<arrayData name="newType"
    encoding="HLAvariableArray" dataType="HLAinteger16BE" cardinality="Dynamic"/>
\end{verbatim}
\end{classdesc}

\begin{classdesc}{HLAfixedRecord}{name, [(fieldName, fieldType)]}
Defines a fixed record.

\begin{verbatim}
recordType = fom.HLAfixedRecord("recordType",
    [('X',fom.HLAinteger16BE), ('Y',fom.HLAfloat32BE)])
\end{verbatim}
is equivalent to the following OMT DIF
\begin{verbatim}
<fixedRecordData name="recordType">
    <field name="X" dataType="HLAinteger16BE"/>
    <field name="Y" dataType="HLAfloat32BE"/>
</fixedRecordData>
\end{verbatim}

The \method{pack} function takes a mapping object.
\begin{verbatim}
encodedValue = fom.recordType.pack({'X':0, 'Y':3.12})
\end{verbatim}
\end{classdesc}

\begin{classdesc}{HLAvariantRecord}{name, (discriminant, dataType), [([enumerators], altName, altType)]}
Defines a variant record.

\begin{verbatim}
recordType = fom.HLAvariantRecord("recordType", ("Axis",AxisEnum),
    [(["TX"],"X",HLAinteger32LE), (["TY",None],"Y",HLAinteger32LE)])
\end{verbatim}
is equivalent to the following OMT DIF
\begin{verbatim}
<variantRecordData name="recordType"
    encoding="HLAvariantRecord" dataType="AxisEnum" discriminant="Axis">
    <alternative enumerator="TX" name="X" dataType="HLAinteger32LE"/>
    <alternative enumerator="TY,HLAother" name="Y" dataType="HLAinteger32LE"/>
</variantRecordData>
\end{verbatim}

The \constant{HLAother} enumerator is represented by the \class{None} value.

The \method{pack} function takes a mapping object.
\begin{verbatim}
encodedValue = fom.recordType.pack({'Axis':AxisEnum.TX, 'X':7.2})
\end{verbatim}
\end{classdesc}

% $Id: module.tex,v 1.3 2008/10/13 17:15:35 gotthardp Exp $
