\documentclass[12pt,a4paper]{howto}
% Download msc.sty from http://satoss.uni.lu/mscpackage
\usepackage{msc}

\newenvironment{hlamsc}[1]{
\begin{figure}[h]
\centering
\pagecolor{white}
\begin{msc}{#1}
\drawframe{no}
\setmscvalues{small}
\setlength{\instdist}{4.5cm}
\setlength{\bottomfootdist}{0cm}
}{
\end{msc}
\textcolor{white}{.} % workaround for the msc.sty clipping error
\end{figure}
}

\title{Using M\&S HLA in Python}
\release{1.0}

\author{Petr Gotthard}
\authoraddress{petr.gotthard at centrum.cz}

\makeindex

\begin{document}
\maketitle

% This makes the Abstract go on a separate page in the HTML version;
% if a copyright notice is used, it should go immediately after this.
%
\ifhtml
\chapter*{Front Matter\label{front}}
\fi

% Copyright statement should go here, if needed.
% ...

% The abstract should be a paragraph or two long.
\begin{abstract}
\noindent
This document describes Python language binding for the Modeling\&Simulation
High Level Architecture (M\&S HLA).
The M\&S HLA is a general purpose architecture for distributed computer
simulation systems. It's defined in [HLA1.3] and [IEEE 1516.1].

The \module{PyHLA} module provides Python wrapper for the C++ HLA API, and
\method{pack}/\method{unpack} methods for IEEE 1516.2 data types.
\end{abstract}

\tableofcontents

\section{Introduction}

The \module{PyHLA} module provides Python language bindings for the
Modeling\&Simulation High Level Architecture (M\&S HLA).
The M\&S HLA is a general purpose architecture for distributed computer
simulation systems. It's defined in [HLA1.3] and [IEEE 1516.1].

The \module{PyHLA} module provides
\begin{itemize}
\item Python language HLA API, that is compliant with the HLA 1.3 standard
(implemented as a Python wrapper for the C++ HLA API),
\item \method{pack}/\method{unpack} methods providing IEEE 1516.2 encoding,
\item \function{HLAuse} function that is able to directly import OMT DIF
datatypes (the XML format described in IEEE 1516.2)
\end{itemize}

The \module{PyHLA} module can be build on a variety of platform/compiler
combinations, including Windows, Linux and Sun Solaris.
The module relies on the Classic Python interpreter (version 2.4 or higher)
and requires a HLA 1.3 compliant RTI with C++ DLC API.

\subsection{Motivation}

The \module{PyHLA} module aims to enable rapid development of HLA federates,
i.e. to simplify the activity 4.3 of FEDEP [IEEE 1516.3].

The HLA standard defines C++ mappings [IEEE 1516.1], but a significant
development effort is necessary to develop a HLA compliant federate in C++.
The development effort can be reduced by using
\begin{enumerate}
\item higher level interfaces, e.g. the Protocol Independent Interface in M\"{A}K VR-Link
\item code generators, e.g. the GENESIS developed by ONERA [GENESIS]
\end{enumerate}

The HLA standard does not cover all aspects.
The C++ API for value encoding [IEEE 1516.2] is not standardized. Every HLA
developer needs to implement the value encoding functions.

Integrating HLA into the Python language may reduce the development and
maintenance effort (compared to C/C++).

The Python language provides several benefits
\begin{enumerate}
\item Dynamic data types. High-level programming language.
\item Interpreted. Easy modifications.
\item Powerful plug-in system.
\item Many plug-ins providing scientific calculations, or geodetic conversions.
\end{enumerate}

\subsection{Installation Instructions}

The \module{PyHLA} module requires
\begin{itemize}
\item Classic Python interpreter 2.4--2.6 (\ulink{CPython}{http://www.python.org/download})
\item HLA1.3 RTI with C++ API (\ulink{CERTI}{http://www.cert.fr/CERTI}, \ulink{M\"{A}K RTI}{http://www.mak.com/products/rti.php})
\end{itemize}

\paragraph{Windows}

A binary installer is provided for the latest CERTI and Python release.

The wizzard will guide you through the installation process. The \module{PyHLA}
module should be installed into the Python \file{site-packages} directory.
When prompted for a destination folder, you should enter something like
\file{C:/Program Files/Python2.5/Lib/site-packages}.

\paragraph{Linux}

The \module{PyHLA} source codes can be obtained from the \ulink{CERTI}{https://savannah.nongnu.org/projects/certi}
applications repository.
\begin{verbatim}
cvs -d:pserver:anonymous@cvs.sv.gnu.org:/sources/certi co applications/PyHLA
\end{verbatim}

To build the \module{PyHLA} module you must provide a path to the Python
\file{site-packages} directory.

If no \makevar{CMAKE_INSTALL_PREFIX} is provided, the path is determined
automatically. The source codes can be compiled and installed using
\begin{verbatim}
cmake .
make install
\end{verbatim}

Alternatively, you may provide the path to the \file{site-packages} directory
in the \makevar{CMAKE_INSTALL_PREFIX} cmake variable.
\begin{verbatim}
cmake . -DCMAKE_INSTALL_PREFIX=/usr/lib/python2.5/site-packages
make install
\end{verbatim}

\subsection{Building the PyHLA documentation}
The \module{PyHLA} documentation may be find online at
\ulink{http://www.nongnu.org/certi/PyHLA/manual}{http://www.nongnu.org/certi/PyHLA/manual}
however you may rebuild the documentation from the PyHLA source itself, in order
to generate either PDF, HTML or DVI format.

The documentation of \module{PyHLA} uses standard Python documentation
system, in particular the python mkhowto script.
For generating the documentation from source you need:
\begin{itemize}
  \item a working python installation with mkhowto script
  \item a working latex compiler
  \item a specific latex style called \texttt{msc.sty} which is used
        to draw Message Sequence Chart in the documentation.
        If you do not have the \texttt{msc.sty} installed you may
        let CMake retrieve the file for you from:        
        \ulink{http://satoss.uni.lu/mscpackage}{http://satoss.uni.lu/mscpackage}.
\end{itemize}

The procedure for generating the documentation is simple,
go the \module{PyHLA} source directory and then:
\begin{verbatim}
make msc.sty
make howto
\end{verbatim}

The \texttt{msc.sty} step is optional and necessary only if you don't have
\texttt{msc.sty} installed.

\subsection{Performance}

Discuss the performance penalty. Use the DFSS/DMAIC methodology.

Design experiments for performance measurements.

Measure performance of Python and plain C++ federates. Use multiple RTI: CERTI and MÄK RTI.

Discuss the results.

\input{rti/module}
\input{omt/module}

\input{PyHLA.ind}
\end{document}

% $Id: PyHLA.tex,v 1.9 2009/07/12 19:30:06 gotthardp Exp $
